Self-tunable approximated explicit MPC: Heat exchanger implementation and analysis

### 2024-04-22 paper/JoPC/response_letter_2nd_round.tex

	 The control performance with Qy = 500 is not quantified, and whether the performance benefits of the proposed method reported in Table 2 still remain is not clarified. Since Qy =500 would give a moderately tuned controller, it can provide a better overall performance than the two boundary controllers chosen with "extreme" tuning.
	}

	\answer{
		We apologize for providing answers that were considered not sufficient. We have carefully revisited our recent manuscript to appropriately address all the Reviewer's comments.

		First, we revisited Table~\ref{tab:control_performance} to evaluate the required control performance criteria for the moderate MPC setup $Q_\mathrm{y} = 500$.
		The values highlighted in blue represent the scenarios when the controller with the investigated setup $Q_\mathrm{y} = 500$ performed as the best one, excluding the self-tunable MPC.

		We recall that Table~\ref{tab:control_performance} summarizes 3 control performance criteria evaluated for 4 operating conditions represented by reference step change. Hence, Table~\ref{tab:control_performance} investigates 12 control cases in total. 
		It can be seen, that the values related to the the setup $Q_\mathrm{y} = 500$ are the best in 5 out of 12 performance criteria, see Table~\ref{tab:control_performance}, blue. On the contrary, the self-tuned controller acquired 10 best values out of 12 performance criteria thanks to the tuning based on the varying operating conditions, see Table~\ref{tab:control_performance}, red.

		When excluding the boundary controllers and comparing only the self-tunable controller with the setup $Q_\mathrm{y} = 500$, the moderate controller outperforms the self-tunable MPC in half of the performance criteria. It happens mostly in tracking the reference value upwards. Nevertheless, when the reference changes downwards, the self-tuned controller becomes significantly beneficial. Therefore, it supports implementing the tunable control technique. To support our statement, the average control performance for all controllers was also elaborated. As the average control performance was the main topic of the Reviewer's comment No.2, these data are analyzed in detail in the answer to the following comment.

@@

		To avoid the unfair evaluation of the relative improvement, we extended Table~\ref{tab:improvement500} such that every performance criterion of the optimal MPC is compared subject to the self-tunable MPC, see Table~\ref{tab:improvement500}. In such way, the base subject to which the relative improvements are calculated, remains the same. Note, that following up on the Reviewer's comment No.1, we also added the evaluation of the controller with the moderate setup, i.e., $Q_\mathrm{y} = 500$ to support the statements from the answer to the previous comment.

		%Regarding the average values of the relative improvements, simple average of the values for every controller across all reference changes would not properly map the overall improvement. For example, let us compare the 79\% improvement in the settling time subject to the MPC setup $Q_\mathrm{y} = 100$ during the third and fourth reference tracking. In the third reference step change, the relative improvement of 79\% corresponds to the settling time difference $31 - 6.5 = 24.5$\,s. In the fourth reference step change, the relative improvement of 79\% corresponds to the settling time difference $90 - 18.5 = 71.5$\,s. Although the values in Table~\ref{tab:improvement500} provide an extended overview of how better or worse the self-tunable controller performed in the specific operating conditions, averaging the values across the reference changes would give a somewhat distorted idea. Therefore, some kind of ``normalization'' would lead to a better vision of the average control performance. Our idea is to first average the values of the control performance criteria for every controller setup across the reference changes, and then calculate the relative improvements from the averaged values, see Table~\ref{tab:average_improvement500}. The symbol $\bar{(.)}$ represents an average value of the corresponding criterion.

		Therefore, based on the explanation of our idea of the average control performance evaluation, at this place we would like to return to the Reviewer's comment No.1--to analyze whether the moderate controller performs better in the overall metric. Let us compare the average relative control improvement of the moderate MPC and the self-tunable approximated MPC. The Table~\ref{tab:improvement500} shows that although in average the SSE criterion corresponding to the self-tunable MPC  was 6\,\% worse, the rest of the criteria improved in average (the maximal overshoot improved by 14\,\% and the settling time improved by 58\,\%). Moreover, we recall that the moderate MPC controller is not advantageous for the decreasing reference step changes. Therefore, considering the proposed tunable technique is beneficial.

@@

		\begin{quote}
			To properly investigate the control results, the control performance was also judged quantitatively using a set of quality criteria. The self-tunable control approach outperformed the conventional control strategy handling just a single controller, i.e., non-tunable controller. \change{In average, the control performance criteria improved compared to the upper and lower boundary MPC respectively as follows: the squared-error-based criterion (SSE) reduced by 23\,\% and 83\,\%, the maximal overshoot/undershoot $\sigma_{\mathrm{max}}$ reduced by 64\,\% and 102\,\%, and the settling time $t_{\epsilon}$ reduced by 12\,\% and 244\,\%.}
		\end{quote}
	}

	\comment{
		Based on the responses to my previous comments, it appears that the applicability of the proposed self-tunable explicit MPC method is limited to "small-scale multivariable systems, with decoupled dynamics"? If so, please mention this clearly in the abstract.
	}

	\answer{
		We incorporated the Reviewer's suggestion into the Abstract as follows:

		\begin{quote}
			The tunable approximated explicit model predictive control (MPC) comes
			with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any intervention from the control engineer during operation in order to retune the controller subject to the changed working conditions. Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. This extension exploits different ranges of the tuning parameter assigned to specified operating conditions. \change{The self-tunable technique is suitable for SISO (single-input and single-output) systems and MIMO (multiple-input and multiple-output) systems with decoupled dynamics.} The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetric behavior of the plant was compensated by tuning the controller’s aggressiveness, as the negative or positive sign of reference change was considered in the tuning procedure. The designed self-tunable controller improved control performance by decreasing sum-of-squared control error, maximal overshoots/undershoots, and settling time compared to the conventional control strategy based on a single (non-tunable) controller.
		\end{quote}

### 2024-04-22 paper/JoPC/main_3rd_round_marked.tex
response letter 2 - almost final?

		\begin{abstract}
			%%max 250 words!			
			The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any interventions of the control engineer during operation in order to \change{retune the controller subject to the changed working conditions.} Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. This extension provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. \textcolor{red}{The self-tunable technique is suitable for SISO (single-input and single-output) systems and is straighforward to implement for MIMO (multiple-input and multiple-output) systems with decoupled dynamics.} The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetric behavior of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning procedure. The designed self-tunable controller improved control performance by decreasing sum-of-squared control error, maximal overshoots/undershoots, and settling time compared to the conventional control strategy based on a single (non-tunable) controller. 
		\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%Research highlights
%		\begin{highlights}
%			 %3-5 highlights, every one should not exceed 85 characters including spaces
%			\item Self-tunable control technique does not need any intervention during control
%			\item Tuning based on operating conditions compensates for the nonlinear plant behavior
%			\item Implementation of the proposed tuning method on a laboratory heat exchanger 
%			\item Control performance improved compared to utilizing explicit MPC with constant setup
%		\end{highlights}

>>>>

		\begin{abstract}
			%%max 250 words!			
			The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any interventions of the control engineer during operation in order to \change{retune the controller subject to the changed working conditions.} Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. This extension provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. \textcolor{red}{The self-tunable technique is suitable for SISO (single-input and single-output) systems and MIMO (multiple-input and multiple-output) systems with decoupled dynamics.} The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetric behavior of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning procedure. The designed self-tunable controller improved control performance by decreasing sum-of-squared control error, maximal overshoots/undershoots, and settling time compared to the conventional control strategy based on a single (non-tunable) controller. 
		\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%Research highlights
%		\begin{highlights}
%			 %3-5 highlights, every one should not exceed 85 characters including spaces
%			\item Self-tunable control technique does not need any intervention during control
%			\item Tuning based on operating conditions compensates for the nonlinear plant behavior
%			\item Implementation of the proposed tuning method on a laboratory heat exchanger 
%			\item Control performance improved compared to utilizing explicit MPC with constant setup
%		\end{highlights}
